# Missing values
```{r, include=FALSE}
source('plot_missing.R')
library(tidyverse)
library(mi)
library(vcd)

knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
```

```{r, results = 'hide'}
tuesdata <- tidytuesdayR::tt_load(2021, week = 38)
spotify <- tuesdata$audio_features
```

```{r, fig.height=8, fig.width=10}
spotify <- spotify %>% rename(
  id = song_id,
  art = performer,
  nam = song,
  gen = spotify_genre,
  trk = spotify_track_id,
  url = spotify_track_preview_url,
  dur = spotify_track_duration_ms,
  exp = spotify_track_explicit,
  alb = spotify_track_album,
  dnc = danceability,
  eng = energy,
  key = key,
  ld = loudness,
  mde = mode,
  sph = speechiness,
  aco = acousticness,
  ins = instrumentalness,
  liv = liveness,
  val = valence,
  tem = tempo,
  sig = time_signature,
  pop = spotify_track_popularity
)

plot_missing(spotify, percent=TRUE)
```

In our exploration of the dataset, we note that the `spotify_track_id` (trk) can uniquely identify songs on Spotify. Whenever `spotify_track_id` is missing, all of the audio features are also missing, as in Patterns 3 and 4 --- we hypothesize that this is the case because Spotify metadata and other measures cannot be accessed without this id. Pattern 10 does not follow this trend, but only occurs in 4 rows total in the entire dataset. 

We also see that missing only `spotify_track_preview_url` (url) is the second most common missing  pattern, at around 34-35%. It is also generally missing from around 50% of our data points; however, this feature is just a link to a 30 second preview of the song. We likely will not find this pattern particularly impactful, as this feature will probably not be utilized in our project. We believe this is feature is MCAR, "Missing Completely At Random." Additionally, we see that the features `song_id` (id), `artist` (art), and `song` (name) do not have any missing data. We believe this is because `song_id` is a unique identifier that is the concatenation of the artist and song columns, and if both song and artist data were missing, the entire row would also be missing because it would be impossible to find the right song.

```{r AG data cleaning, echo=FALSE}
rawdata <- readr::read_csv("billboard_spotify_cleaned.csv")
rawdata$spotify_genre_clean <- as.factor(rawdata$spotify_genre_clean)
rawdata$week_id <- as.Date(rawdata$week_id, format="%m/%d/%Y")

data <- rawdata
data$Missing <-ifelse(rowSums(is.na(data)) >= 1, TRUE, FALSE)
# sort by (week?), song, then artist
data <- data[order(data$performer,
                   data$song,
                   data$week_id),]
summdata <- data %>% 
  group_by(song, performer) %>%
  summarise(first_week = min(week_id),
            last_week = max(week_id),
            total_peak = min(peak_position),
            total_instances = max(instance),
            total_weeks=n()) %>% # found to be more accurate than max(weeks_on_chart)
  ungroup()

uniqdata <- data %>%
  select(c("spotify_genre_clean", "song", "performer",
           "key", "mode", "tempo", "time_signature",
           "danceability", "energy", "loudness", 
           "speechiness", "acousticness", "instrumentalness",
           "liveness", "valence",
           "Missing", # custom column
           "spotify_genre", "spotify_track_popularity")) %>% 
  # keeping original genre to check
  unique() %>%
  inner_join(summdata, by=c("song", "performer"))
```

```{r AG Missingness over time, echo=FALSE}
uniqdata %>%
  mutate(year = as.numeric(format(last_week, format="%Y"))) %>%
  group_by(year) %>%
  summarise(Total = sum(Missing)/n(),
            Genre = sum(is.na(spotify_genre_clean))/n()) %>%
  # need percentages due to non-standard number of songs/year
  ungroup() %>%
  pivot_longer(cols=c(Total, Genre),
               names_to = "type",
               values_to = "perc") %>% 
  mutate(perc = 100*perc,
         type = as.factor(type)) %>%
  ggplot(mapping=aes(x=year, y=perc, 
                     color=fct_relevel(type, "Total", "Genre"), 
                     group=fct_relevel(type, "Total", "Genre"))) +
  geom_line() +
  scale_color_discrete(name="Name type") +
  scale_x_continuous(limits = c(1958, 2021)) + 
  scale_y_continuous(limits = c(0, 50)) +
  labs(x="Year", y="% Missing", 
       title="Missingness by Year") +
  theme(plot.title = element_text(size = 16, hjust=0.5),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
```

```{r AG Missingness featured, echo=FALSE}
temp <- uniqdata %>%
  mutate(Featuring = str_detect(performer, regex('[:space:]ft[:space:]', ignore_case=TRUE)) |
           str_detect(performer, regex('[:space:]ft.[:space:]', ignore_case=TRUE)) |
           str_detect(performer, regex('[:space:]feat[:space:]', ignore_case=TRUE)) |
           str_detect(performer, regex('[:space:]feat.[:space:]', ignore_case=TRUE)) |
           str_detect(performer, regex('[:space:]featuring[:space:]', ignore_case=TRUE)) |
           str_detect(song, regex('[:space:]featuring[:space:]', ignore_case=TRUE))) %>%
  # almost all are of the form "[:space:]featuring[:space:]" in performer
  # many features are actually recorded in the form "[artist] & [ft. artist]"
  # sometimes "with" or "x" or other choices, as well
  # but these cannot be easily distinguished from bands with "&" in their names
  group_by(Featuring, Missing) %>%
  summarise(Freq = n()) %>% # vcd::mosaic convention
  ungroup() %>% # relevel factors 
  mutate(Featuring = recode(as.factor(Featuring), "TRUE" = "True", "FALSE" = "False"),
         Missing = recode(as.factor(Missing), "TRUE" = "True", "FALSE" = "False"))


mosaic(Missing~Featuring, direction = c("v", "h"), temp, 
       highlighting_fill = c("#FDAE61","#D53E4F"),
       main = "Missingness in Songs w/ Featured Artists",
       gp_varnames = gpar(fontsize = 14, fontface = 1),
       gp_labels = gpar(fontsize = 12))
```

  Once we expanded our dataset to join the Billboard ranking and Spotify metadata information, we wanted to expand our understanding of the missingness patterns in our dataset and how they might be related with the dataset creators’ ability to access certain information using the Spotify API.
  
  Our first hypothesis was that older songs might have a higher percentage of rows with missing data. In order to investigate this, we marked every row with at least one column with an NA as “missing” and calculated the percent of unique songs (assuming each song in the raw data with missing values was missing that same data for all of its associated entries for each week it was in the Top 100) from each year that had incomplete data. We also separately calculated a `missing` percentage for the `spotify_genre` column in particular, as our initial observation that many songs with featured artists did not have genre information, and we thought there might be a different pattern to missing spotify information. Although this second hypothesis does not seem to be true, we can see that our intuition for our first one was generally correct and that it could be helpful to remove some of the earlier decades from our dataset in certain analyses.
  
  Additionally, because the Spotify join was on the `song` and `performer` columns, songs that included featured artists--which are often noted in various nonstandard ways, from multiple abbreviations of “featuring” to uses of “&” and other separating characters--would prevent us from properly identifying a song and accessing its features. If genre information was missing, we believed all other Spotify song features would also be missing. Thus, we searched for all songs featuring additional artists as best as possible, and plotted the proportions of missingness between songs with and without. Our results here do not seem to give evidence supporting our hypothesis, and we also found that the randomness to genre’s missing pattern is due to its existence at both the album level and individual track level (while song information exists only per track). 