# Results

```{r, echo=FALSE}
data <- readr::read_csv("billboard_spotify_cleaned_.csv")
```

##What features define song popularity through the decades?
```{r, echo = FALSE}
top_5<- data %>% 
          filter(peak_position <= 5) %>%
          mutate(date = as.Date(week_id,  "%m/%d/%Y"),
                 song_id = paste(song, performer, sep="_"),
                 year = year(date),
                 decade = paste(as.character(floor(year/10) *10),"'s", sep="")) %>%
          filter(year >= 1960) %>%
          group_by(song_id, instance) %>%   #including instance: A song will show up mult time/year
          mutate(abs_peak_position = min(peak_position),
                 max_weeks_on_chart = max(weeks_on_chart)) %>%
          distinct(abs_peak_position, .keep_all = TRUE) %>%
          ungroup()

top_5 %>% 
  select(song_id, song_features, decade) %>% 
  mutate(loudness = loudness/-60 ) %>%
  pivot_longer(song_features, values_to = 'values', names_to='features', values_drop_na = TRUE) %>%
  ggplot(aes(x=values, y = features, fill = features)) +
  geom_density_ridges(alpha = 0.5) +
  scale_fill_cyclical(name = "Cycle", values = c("lightblue", "blue"))+
  scale_color_cyclical(name = "Cycle", values = c("lightblue", "blue")) +
  ggtitle("Top 5 Songs' features") + xlab('Value') + ylab('Song Features') + facet_wrap(~decade)


```

Some skew toward higher valence (more positive), low speechiness (presence of words). Values above 0.66 describes tracks that are mostly spoken words. 0.33-0.66 represents tracks with music and speech (like rap). loudness correlates to physical strength, scaled to the range between -60 to 0 db. Liveness greater than 0.8 indicates a strong likelihood that song is live. Instrumentalness (closer to 1 = more likely no vocals). 

```{r, echo = FALSE, fig.height = 7, fig.width = 12}
par_cord_kq <-top_5 %>%
            filter(abs_peak_position == 1, (decade == "1960's"| decade =="2010's")) %>%
            mutate(loudness = loudness/-60) %>%
            select(decade, mode, song_features)


ggparcoord(data = par_cord_kq, columns = 2:10, 
           scale = 'uniminmax', 
           groupColumn = 'decade',
           alphaLines = 1, splineFactor = 7,
           title = "Features of Songs that Peaked at the No.1 on Charts") +
  xlab('Song Feature')



```




```{r, echo = FALSE, fig.height = 15, fig.width=15}
genre_5 <- top_5 %>%
            subset(spotify_genre_clean != 'NA') %>%
            rename(Genre = spotify_genre_clean,
                   Decade = decade)
genre_names <- distinct(genre_5, Genre)[1]
num_colors <- count(genre_names)
fills18 <- colorRampPalette(brewer.pal(num_colors, 'Spectral'))

genre_5$Genre <- fct_rev(fct_infreq(genre_5$Genre))
genre_5$Genre <- fct_relevel(genre_5$Genre, "trap", "jazz", "indie", "k-pop","punk", "house", "blues", "edm", "funk", "country", "r&b", "other")

mosaic(Genre ~ Decade, 
       direction = c("v", "h"),
       highlighting_fill = fills18,
       data = genre_5,
       labeling= labeling_border(rot_labels = c(0,0,0,0)),
       spacing = vcd::spacing_equal(sp = unit(0, "lines")))


```
Notably, there is a shift (and creation) of some genres becoming more or less popular during certain decades. For example, it's evident that Blues were popular in the 19050s but have progressively become less popular in subsequent decades whereas Country became more popular, peaking in the 70s before losing relative popularity. Some genres such as edm, k-pop, hip-hop and rap didn't even exist in earlier decades, and did not show any popularity until the last 10-20 years.


##What artists and songs have trended since 2000?
```{r data cleaning}

data_2000 <- data %>%
              mutate(week_id = as.Date(data$week_id, format = '%m/%d/%Y')) %>%
              filter(week_id >= as.Date("2000-1-1"))

data_filtered <- data_2000[,1:9]
```

```{r top 50 songs by # of weeks on charts}
top_song <- data_filtered %>% 
  group_by(song, performer) %>% 
  slice(which.max(weeks_on_chart)) %>% 
  arrange(desc(weeks_on_chart)) %>% 
  head(50)


ggplot(top_song, aes(x = weeks_on_chart, y = reorder(song, weeks_on_chart), color=spotify_genre)) +
  geom_point()


```

Insight:


```{r top song by # of weeks at #1}
top_song_1 <- data_filtered %>% 
  group_by(song) %>% 
  filter(week_position == 1) %>% 
  add_count() %>% 
  slice(which.max(week_id)) %>% 
  arrange(desc(n)) %>% 
  head(50)

ggplot(top_song_1, aes(x = n, y = reorder(song, n), color = spotify_genre)) +
  geom_point()


```

Insight:


```{r top artists by number of songs on chart}
top_artists <- data_filtered %>% 
  group_by(performer, song) %>% 
  slice(which.max(week_id)) %>% 
  group_by(performer) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(50)

ggplot(top_artists, aes(x = n, y = reorder(performer, n))) +
  geom_point()
```

Insight:


```{r top artists by number of #1 songs on chart}
top_artists_1 <- data_filtered %>% 
  group_by(performer, song) %>% 
  filter(week_position==1) %>% 
  slice(which.max(week_id)) %>% 
  group_by(performer) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(35)

ggplot(top_artists_1, aes(x = n, y = reorder(performer, n))) +
  geom_point()
```

Insight:

##What defines a genre and what makes k-pop so unique?

```{r AG Uniq songs}
rawdata <- readr::read_csv("billboard_spotify_cleaned_kpop.csv")
rawdata$spotify_genre_clean <- as.factor(rawdata$spotify_genre_clean)
rawdata$week_id <- as.Date(rawdata$week_id, format="%m/%d/%Y")

data <- rawdata
data$missing <-ifelse(rowSums(is.na(data)) >= 1, TRUE, FALSE)
# sort by (week?), song, then artist
data <- data[order(data$performer,
                   data$song,
                   data$week_id),]
summdata <- data %>% 
  group_by(song, performer) %>%
  summarise(first_week = min(week_id),
            last_week = max(week_id),
            total_peak = min(peak_position),
            total_instances = max(instance),
            total_weeks=n()) %>% # found to be more accurate than max(weeks_on_chart)
  ungroup()

uniqdata <- data %>%
  select(c("spotify_genre_clean", "song", "performer",
           "key", "mode", "tempo", "time_signature",
           "danceability", "energy", "loudness", 
           "speechiness", "acousticness", "instrumentalness",
           "liveness", "valence",
           "missing", # custom column
           "spotify_genre", "spotify_track_popularity")) %>% 
  # keeping original genre to check
  unique() %>%
  inner_join(summdata, by=c("song", "performer"))

# manually alter 'CRUSH' to non-kpop
uniqdata[which(uniqdata$performer=="CRUSH"),1] <- "other"

first <- uniqdata %>%
  mutate(year = as.numeric(format(first_week, format="%Y"))) %>%
  group_by(year) %>%
  summarise(First = n()) %>%
  ungroup()

last <- uniqdata %>%
  mutate(year = as.numeric(format(last_week, format="%Y"))) %>%
  group_by(year) %>%
  summarise(Last = n()) %>%
  ungroup() 

cbind(first, last[,-1]) %>%
  pivot_longer(cols=c(First, Last),
               names_to = "Year_bin",
               values_to = "Freq") %>%
  ggplot(mapping=aes(x=year, y=Freq,
                     color=Year_bin,
                     group=Year_bin)) +
  geom_line() +
  scale_color_discrete(name="Year Charted") +
  scale_x_continuous(limits = c(1958, 2021)) +
  scale_y_continuous(limits = c(0, 800)) +
  labs(x="Year", y="Frequency",
       title="Number of unique songs in the Top 100")  +
  theme_gray()
# FIX FCT_REORDER TO MATCH FIGURE
# add annotated lines of best fit
```

Insight:

```{r AG Biplot of auditory features}
fulldata <- na.omit(uniqdata)

# histogram of song popularity
fulldata %>%
  pivot_longer(cols=c(total_peak,
                      total_instances,
                      total_weeks,
                      spotify_track_popularity),
               names_to = "feature",
               values_to = "value") %>%
  ggplot(mapping=aes(value)) +
  geom_histogram() +
  facet_wrap(~feature, scales="free") +
  labs(x="Value", y="Count",
       title="Spreads of song popularity measures") +
  theme_gray()

# histogram of all features
fulldata %>%
  pivot_longer(cols=key:valence,
               names_to = "feature",
               values_to = "value") %>%
  ggplot(mapping=aes(value)) +
  geom_histogram() +
  facet_wrap(~feature, scales="free") +
  labs(x="Value", y="Count",
       title="Spreads of musical features") +
  theme_gray()
# thus choose acousticness (12), danceability (8), energy (9), 
# liveness (14), loudness (10), tempo (6), and valence (15)
# removed speechiness (11) after looking at the pca summary
featcols <- c(6,8,9,10,11,12,14,15)

pcs <- prcomp(fulldata[,featcols], scale.=TRUE)
expvar <- 100*(pcs$sdev^2)/sum(pcs$sdev^2)

# https://stackoverflow.com/questions/6578355/plotting-pca-biplot-with-ggplot2
datax <- data.frame(pcs$x)
datapc <- data.frame(varnames=rownames(pcs$rotation), pcs$rotation)
mult <- min(
  (max(datax[,"PC2"]) - min(datax[,"PC2"])/(max(datax[,"PC2"])-min(datax[,"PC2"]))),
  (max(datax[,"PC1"]) - min(datax[,"PC1"])/(max(datax[,"PC1"])-min(datax[,"PC1"])))
)
datapc <- transform(datapc,
                    v1 = .8 * mult * (get("PC1")),
                    v2 = .8 * mult * (get("PC2")))

temp <- cbind(data.frame(pcs$x), fulldata) %>%
  filter(total_peak < 50) 

temp %>% 
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(aes(color = spotify_genre_clean, 
                 alpha = 0.01, stroke = 0)) +
  geom_point(subset(temp, spotify_genre_clean == 'kpop'),
             mapping=aes(color = spotify_genre_clean)) +
  geom_text(data=datapc, aes(x=v1, y=v2, label=varnames),
            nudge_x=ifelse(datapc$v1 > 0, 0.3, -0.3),
            nudge_y=ifelse(datapc$v2 > 0, 0.4, -0.4), 
            angle=0,
            size=4, color="black") +
  geom_segment(data=datapc, mapping=aes(x=0, y=0, xend=v1, yend=v2), 
               arrow=arrow(length=unit(0.2,"cm")), 
               alpha=0.75, color="black") +
  scale_alpha(guide="none") +
  scale_size(guide="none") +
  scale_color_discrete(name="Genre") +
  labs(x=paste("PC 1 (",format(expvar[1], digits=3),"%)"),
       y=paste("PC 2 (",format(expvar[2], digits=3),"%)"),
       title="Biplot of musical features") +
  theme_gray()
```

Insight:

```{r AG parcoords kpop}
# REQUIRES MOST TWEAKING
fulldata %>%
  mutate(year = as.numeric(format(first_week, format="%Y"))) %>%
  filter(((year >= 2009) & (total_peak <= 5)) | 
           (spotify_genre_clean == "kpop")) %>% 
  ggparcoord(columns = featcols, scale="uniminmax", alphaLines=0.1, splineFactor=10,
             groupColumn = 'spotify_genre_clean') +
  labs(title="Auditory features", x="Auditory features", y="Count") +
  theme(legend.position = "none") 
```

Insight: